Wrattler: \textnormal{A platform for AI-assisted data science}

### 2018-03-16 report/paper.tex
update wrt typos

%\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
%\acmYear{2017}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\setcopyright{none}
\bibliographystyle{ACM-Reference-Format}

@@

\author{AIDA Team, The Alan Turing Institute}
%\affiliation{
%  \institution{The Alan Turing Institute}
%  \country{London, United Kingdom}
%}
%\email{tomas@tomasp.net}

@@

\paragraph{Big data is big,}
so the analyst doesn't understand it all.*
Even if a data set is small enough fits on one computer, it's still too large to fit in
an analyst's working memory.
But this means every analysis is huanted
by the spectre of lurking, potentially
unknown data quality  issues.
This also makes it more difficult to do data fusion, because there may be corner cases that make it more difficult to join two disparate data sources than expected.

@@

The data store provides a way for persistently storing data and enables communication between
individual Wrattler components. Data store keeps external data files imported into Wrattler,
as well as computed data frames (and, potentially, other data structures). It is immutable
and keeps past versions of data to allow efficient state rollback. For persistency and versioning,
data store also serializes and stores multiple versions of the dependency graph.

The data store supports a mechanism for annotating data frames with additional semantic information, such as:
%
\begin{itemize}
\item[--] Columns can be annotated with (and stored as) primitive data type such as date or floating-point number.
\vspace{0.25em}
\item[--] Columns (or a combination) can be annotated with semantic annotation such as geo-location or address.
\vspace{-0.85em}
\item[--] Columns, rows and cells of the data frame can be annotated with other metadata such as provenance.
\end{itemize}

@@

\begin{figure}
\begin{equation*}
\begin{array}{l}
\textit{\small{In the first block, we retrieve data from a government web site.}}\\[-0.1em]
\textit{\small{We use Python to download the data and parse it as a CSV file.}}\\[0.3em]
\quad\ident{url}=\ident{open}(\str{http://data.gov.uk/\ldots/bb2014.csv})\\
\quad\ident{bbRaw}=\ident{csv.reader}(\ident{url})\\[0.2em]
\textit{\small{In the next block, we used an AI assistant to assist with cleaning}}\\[-0.1em]
\textit{\small{and created a script that marks \#!NULL values as missing:}}\\[0.3em]
\quad\kvd{let}~\ident{bbClean} = \ident{bbRaw}.\ident{missing}.\ident{as\_missing}(\str{\#!NULL})\\[0.3em]
\textit{\small{Finally, we use JavaScript to create a histogram of upload speeds:}}\\[0.3em]
\quad\ident{vega}.\ident{hsit}(\ident{bbClean}, \str{Upload speed})
\end{array}
\end{equation*}
\vspace{-0.5em}
\caption{\small{A small notebook consisting of three cells. The first Python cell exports
\ident{bbRaw}; the second Wrattler cell uses it and exports \ident{bbClean} and the
JavaScript cell visualizes \ident{bbClean}.}}
\label{fig:notebook}
\end{figure}

@@

\caption{\small{Dependency graph for the notebook in Figure~\ref{fig:notebook}.
Data acquisition is represented as a single Python cell. Data cleaning is understood
by Wrattler and is the graph has a node for each call and constant. Hexagons represent
data frames shared by cell and stored in the data store.}}
\label{fig:deps}
\vspace{-0.5em}
\end{figure}

@@

\paragraph{Performing analytics.}
An increasing amount of attention is being
given to the AutoML challenge \cite{guyon_review_2016}, which focuses
on automatically choosing a classification
methods, feature representations and hyperparameters
for a given classification task, without the need for human intervention.

\paragraph{Debugging and interpreting predictive analytics.}
Every time a new machine learning model is trained, the first question
an analyst has is often why its accuracy is not better. But debugging
predictive analytics is notoriously difficult, and the remedies
indirect and expensive, e.g., labelling more data, adding more
features, using a more complex model. Debugging and improving models
is a rich area for potential assistants.  There is potentially low
hanging fruit available in how to aid the tasks of error analysis,
exploring the predictions of the model on validation data. Some early
work in this direction is \cite{saleema}.  There is a rich and growing
literature on interpreting and explaining the predictions of a model
\cite{lipton:mythos,doshi-velez17,ribiero2016lime,darksight}.  Two
potential directions from are debugging by using "training set
blame": Given an error in the validation data, what examples from the
training set caused me to make that error? This might build on the
work of \cite{percy}.  Another potential avenue is explaining
differences between predictions: why was instance 42 treated
differently than instance 24601, even though they appear highly
similar to a human?  }

@@

In this section, we focus on the initial functionality that is
required from the data store.  We focus on the public interface
exposed by the data store, but we do not discuss the technology used
for data storage -- this can be decided during the development.

Data frames stored in the data store can be annotated with
metadata. Such metadata can communicate information between multiple
AI assistants. For example, first AI assistant can infer that a pair
of columns represent GPS lat/lon coordinates and produce a data frame
with annotations specifying this information. A second AI assistant
can then use this information and automatically provide a data
visualization in the form of map. In the rest of this section, we
discuss the structure of those annotations.

@@

\bibliographystyle{plain}
\bibliography{paper}
\end{document}

>>>>

%\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
%\acmYear{2017}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\setcopyright{none}

@@

\author{AIDA Team, The Alan Turing Institute \\ 16 March 2018}

%\affiliation{
%  \institution{The Alan Turing Institute}
%  \country{London, United Kingdom}
%}
%\email{tomas@tomasp.net}

@@

\paragraph{Big data is big,}
so the analyst doesn't understand it all.
Even if a data set is small enough to fit on one computer, it is still too large to fit in
an analyst's working memory.
But this means every analysis is huanted
by the spectre of lurking, potentially
unknown data quality  issues.
This also makes it more difficult to do data fusion, because there may be corner cases that make it more difficult to join two disparate data sources than expected.

@@

The data store provides a way for persistently storing data and enables communication between
individual Wrattler components. The data store keeps external data files that have been
imported into Wrattler,
as well as computed data frames (and, potentially, other data structures). It is immutable
and keeps past versions of data to allow efficient state rollback. For persistency and versioning,
data store also serializes and stores multiple versions of the dependency graph.

The data store supports a mechanism for annotating data frames with additional semantic information, such as:
%
\begin{itemize}
\item[--] Columns can be annotated with (and stored as) primitive data type such as date or floating-point number.
\item[--] Columns (or a combination) can be annotated with semantic annotation such as geo-location or address.
\item[--] Columns, rows and cells of the data frame can be annotated with other metadata such as provenance.
\end{itemize}

@@

\begin{figure}
\begin{equation*}
\begin{array}{l}
\textit{\small{In the first block, we retrieve data from a government web site.}}\\[-0.1em]
\textit{\small{We use Python to download the data and parse it as a CSV file.}}\\[0.3em]
\quad\ident{url}=\ident{open}(\str{http://data.gov.uk/\ldots/bb2014.csv})\\
\quad\ident{bbRaw}=\ident{csv.reader}(\ident{url})\\[0.2em]
\textit{\small{In the next block, we used an AI assistant to assist with cleaning}}\\[-0.1em]
\textit{\small{and created a script that marks \#!NULL values as missing:}}\\[0.3em]
\quad\kvd{let}~\ident{bbClean} = \ident{bbRaw}.\ident{missing}.\ident{as\_missing}(\str{\#!NULL})\\[0.3em]
\textit{\small{Finally, we use JavaScript to create a histogram of upload speeds:}}\\[0.3em]
\quad\ident{vega}.\ident{hist}(\ident{bbClean}, \str{Upload speed})
\end{array}
\end{equation*}
\vspace{-0.5em}
\caption{\small{A small notebook consisting of three cells. The first Python cell exports
\ident{bbRaw}; the second Wrattler cell uses it and exports \ident{bbClean} and the
JavaScript cell visualizes \ident{bbClean}.}}
\label{fig:notebook}
\end{figure}

@@

\caption{\small{Dependency graph for the notebook in Figure~\ref{fig:notebook}.
Data acquisition is represented as a single Python cell. Data cleaning is understood
by Wrattler and is the graph has a node for each call and constant. Hexagons represent
data frames shared by multiple cells and stored in the data store.}}
\label{fig:deps}
\vspace{-0.5em}
\end{figure}

@@

\paragraph{Performing analytics.}
An increasing amount of attention is being
given to the AutoML challenge \cite{guyon_review_2016}, which focuses
on automatically choosing classification
methods, feature representations and hyperparameters
for a given classification task, without the need for human intervention.

\paragraph{Debugging and interpreting predictive analytics.}
Every time a new machine learning model is trained, the first question
an analyst has is often why its accuracy is not better. But debugging
predictive analytics is notoriously difficult, and the remedies
indirect and expensive, e.g., labelling more data, adding more
features, using a more complex model. Debugging and improving models
is a rich area for potential assistants.  There is potentially low
hanging fruit available in how to aid the tasks of error analysis,
exploring the predictions of the model on validation data. Some early
work in this direction is that of \citet{ren2016squares}.  There is a rich and growing
literature on interpreting and explaining the predictions of a model
\cite{lipton:mythos,doshi-velez17,Ribeiro2016lime,darksight}.  Two
potential directions from are debugging by using "training set
blame": Given an error in the validation data, what examples from the
training set caused me to make that error? This might build on the
work of \citet{koh2017understanding}.  Another potential avenue is explaining
differences between predictions: why was instance 42 treated
differently than instance 24601, even though they appear highly
similar to a human?  }

@@

In this section, we focus on the initial functionality that is
required from the data store.  We focus on the public interface
exposed by the data store, but we do not discuss the technology used
for data storage; this is to be decided during the development.

Data frames stored in the data store can be annotated with
metadata. Such metadata can communicate information between multiple
AI assistants. For example, suppose that the first AI assistant  infers that a pair
of columns represent GPS latitude/longitude coordinates and produces a data frame
with annotations specifying this information. A second AI assistant
can then use this information and automatically provide a data
visualization in the form of map. In the rest of this section, we
discuss the structure of those annotations.

@@

\bibliographystyle{plainnat}
\bibliography{paper}
\end{document}
